FHJ: A Formal Model for Hierarchical Dispatching and Overriding

### 2018-05-08 ECOOP18/sections/Discussion.tex
temp

\section{Implementation \& Discussion}\label{sec:discussion}
In this section, we will talk about the prototype implementation and discuss the design space and reflect about some of the design decisions of our work. We relate our language to traits, Java interfaces as well as other languages. Furthermore, we discuss ways to improve our work.

>>>>

\section{Discussion}\label{sec:discussion}
In this section, we will talk about the prototype implementation and discuss the design space and reflect about some of the design decisions of our work. We relate our language to traits, Java interfaces as well as other languages. Furthermore, we discuss ways to improve our work.

### 2018-05-08 ECOOP18/sections/Formalization.tex
temp

\subsubsection{Interfaces}
In order to achieve multiple inheritance, an interface can have a set of 
parent interfaces, where such a set can be empty. The interface declaration $\interface{I}{I}{M}$ introduces an interface named $I$ with parent interfaces $\overline{I}$ and a suite of methods $\overline{M}$. The methods of $I$ may either override methods that are already defined in $\overline{I}$ or add new functionality special to $I$, we will illustrate this in more detail later.

>>>>

\subsubsection{Interfaces}
\yanlin{Bruno: acyclic, clarified herer.}
In order to achieve multiple inheritance, an interface (which is acyclic, by deafault) can have a set of 
parent interfaces, where such a set can be empty. The interface declaration $\interface{I}{I}{M}$ introduces an interface named $I$ with parent interfaces $\overline{I}$ and a suite of methods $\overline{M}$. The methods of $I$ may either override methods that are already defined in $\overline{I}$ or add new functionality special to $I$, we will illustrate this in more detail later.

### 2018-05-08 ECOOP18/sections/Introduction.tex
temp

To present hierarchical overriding and dispatching, we introduce a
formalized model \MIM{} in Section~\ref{sec:formalization} based on
Featherweight Java~\cite{Igarashi01FJ}, together with theorems and
proofs for type soundness. 
In summary, our contributions are:
\begin{itemize}
	\item \textbf{A formalization of the hierarchical dispatching algorithm} that integrates both the static type and dynamic type for method dispatch, and 
	ensures unambiguity as well as extensibility in the presence
        of unintentional method conflicts.
	\item \textbf{Hierarchical overriding:} a novel notion that allows
          methods to override individual branches of the class hierarchy.
	\item \textbf{\name:} a formalized model based on
          Featherweight Java, supporting the above features. 
          We provide the static and dynamic semantics and prove the
          type soundness of the model.
	\item \textbf{Prototype implementation\footnote{The implementation is available at \url{https://github.com/YanlinWang/MIM/tree/master/Calculus}}:} a
          simple implementation of \MIM{} interpreter in Scala. The implementation can type-check and run variants
          of all the examples shown in this paper.
\end{itemize}

>>>>

To present hierarchical overriding and dispatching, we introduce a
formalized model \MIM{} in Section~\ref{sec:formalization} based on
Featherweight Java~\cite{Igarashi01FJ}, together with theorems and
proofs for type soundness.

\yanlin{Bruno.}
In terms of implementation, we have a prototype implementation of \name{} interpreter in Scala. The implementation validates all the examples presented in the paper. The implementation is done in 
Scala, strictly following the type-checking rules and small-step semantics presented in the paper. One nice feature of the implementation is that it can show the detailed step-by-step evaluation of the program, which is convenient for understanding and debugging programs \& semantics. The final model of \name{}
is actually a bit different from our initial models. This is partly because in
the process of implementing and proving theorems we found errors and modified the earlier models, creating a positive feedback circle.

In summary, our contributions are:
\begin{itemize}
	\item \textbf{A formalization of the hierarchical dispatching algorithm} that integrates both the static type and dynamic type for method dispatch, and 
	ensures unambiguity as well as extensibility in the presence
        of unintentional method conflicts.
	\item \textbf{Hierarchical overriding:} a novel notion that allows
          methods to override individual branches of the class hierarchy.
	\item \textbf{\name:} a formalized model based on
          Featherweight Java, supporting the above features. 
          We provide the static and dynamic semantics and prove the
          type soundness of the model.
	\item \textbf{Prototype implementation\footnote{The implementation is available at \url{https://github.com/YanlinWang/MIM/tree/master/Calculus}}:} a
          simple implementation of \MIM{} interpreter in Scala. The implementation can type-check and run variants
          of all the examples shown in this paper.
\end{itemize}

### 2018-05-08 ECOOP18/sections/Overview.tex
temp

\subsubsection{Problem and Possible Workarounds} The essential problem is
how to resolve unintentional method conflicts and invoke the
conflicting methods separately without ambiguity. To tackle this problem, there are several other workarounds
that come to our mind. We briefly discuss those potential fixes and
workarounds next:
\begin{itemize}
  \item \textit{I. Delegation.} As an alternative to multiple inheritance,
  delegation can be used by introducing two fields (or field methods) with the \yanlin{"field methods"?}
  \lstinline|Drawable| type and \lstinline|Deck| type,
  respectively. Although it avoids method conflicts, it is known
  that using delegation makes it hard to correctly maintain
  self-references in an extensible system and also
  introduces a lot of boilerplate code.

@@

\subsubsection{Potential solutions/workarounds in existing languages} Unfortunately in all languages we know of 
(including C++), the existing approaches are unsatisfactory.
One direction is to simply avoid this issue, by putting overriding before inheritance. For example, as shown in Figure~\ref{fig:drawablesafedeck}(middle), we define a new component \lstinline|DrawableRect| that extends \lstinline|Drawable|, which simply draws the deck as a rectangle, and modifies the hierarchy:
\begin{lstlisting}
interface DrawableRect extends Drawable {
  void draw() {
    JFrame frame = new JFrame("Canvas");
    frame.setSize(600, 600);
    frame.getContentPane().setBackground(Color.red);
    frame.getContentPane().add(new Square(10,10,100,100)); ...
  }
}
interface DrawableSafeDeck extends DrawableRect, SafeDeck {}
\end{lstlisting}
This workaround seems to work, but there are severe issues:
\begin{itemize}
	\item It changes the hierarchy and existing code, hence breaks the modularity.
	\item Separate overriding is required to come after the \wordfork{} 
          inheritance, especially when the implementation needs functionality from both parents. In the above code, we have assumed
	that the overriding is unrelated to \lstinline|Deck|. But when the drawing relies on some information of the \lstinline|Deck|
	object, we have to either introduce field methods for delegation or change the signature of \lstinline|draw| to take a parameter.
	Either way introduces unnecessary complexity and affects extensibility.
\end{itemize}

>>>>

\subsubsection{Problem and Possible Workarounds} The essential problem is
how to resolve unintentional method conflicts and invoke the
conflicting methods separately without ambiguity. To tackle this problem, there are several other workarounds
that come to our mind. We briefly discuss those potential fixes and
workarounds next:
\begin{itemize}
  \item \textit{I. Delegation.} As an alternative to multiple inheritance,
  delegation can be used by introducing two fields (or field methods) with the 
  \lstinline|Drawable| type and \lstinline|Deck| type,
  respectively. Although it avoids method conflicts, it is known
  that using delegation makes it hard to correctly maintain
  self-references in an extensible system and also
  introduces a lot of boilerplate code.

@@

\subsubsection{Potential solutions/workarounds in existing languages} 
\yanlin{Bruno: one reviewer says S2.3.1 is "less clear", could you check again?}Unfortunately in all languages we know of 
(including C++), the existing approaches are unsatisfactory.
One direction is to simply avoid this issue, by putting overriding before inheritance. For example, as shown in Figure~\ref{fig:drawablesafedeck}(middle), we define a new component \lstinline|DrawableRect| that extends \lstinline|Drawable|, which simply draws the deck as a rectangle, and modifies the hierarchy:
\begin{lstlisting}
interface DrawableRect extends Drawable {
  void draw() {
    JFrame frame = new JFrame("Canvas");
    frame.setSize(600, 600);
    frame.getContentPane().setBackground(Color.red);
    frame.getContentPane().add(new Square(10,10,100,100)); ...
  }
}
interface DrawableSafeDeck extends DrawableRect, SafeDeck {}
\end{lstlisting}
This workaround seems to work, but there are severe issues:
\begin{itemize}
	\item It changes the hierarchy and existing code, hence breaks the modularity.
	\item Separate overriding is required to come after the \wordfork{} 
          inheritance, especially when the implementation needs functionality from both parents. In the above code, we have assumed
	that the overriding is unrelated to \lstinline|Deck|. But when the drawing relies on some information of the \lstinline|Deck|
	object, we have to either introduce field methods for delegation or change the signature of \lstinline|draw| to take a parameter.
	Either way introduces unnecessary complexity and affects extensibility.
\end{itemize}
